---
title: "ST558 Project 2"
author: "Yu Jiang"
date: "7/3/2020"
params:
  weekday: weekday_is_monday
output: 
  rmarkdown::github_document
---

# Introduction

## Describe the Data

Many articles have been published by Mashable over two years from seventh Jan, 2013 and this original dataset can be found [this website](https://archive.ics.uci.edu/ml/datasets/Online+News+Popularity). 

There are 61 variables in total from the dataset above: 58 predictive attributes, 2 non-predictive and 1 goal field. More details about used data will be discussed later.

## the Purpose of Analysis

We are interested in predicting the popularity of a given article and thus we are going to  treat *shares*, an index of the article popularity, as the response of models and some other variables as predictors. 

After fitting the linear and ensemble models respectively, we can choose the best two different types of models by comparing some numeric values, like AIC and BIC, and then we can make a conclusion about *shares*.

## Methods

First, we can slice data into two sets, a training set (70% of the data) and the test set (30% of the data) and then we are going to start the model fit.

Second, we are going to consider two types of models.

For the linear regression model, We will try to predict popularity in two ways. The first target is as it is recorded, as the number of times the articles is shared over the period. The second target is a binary variable which discretizes the above. Following Fernandes et al. (section 3.1), an article is considered popular if it exceeds 1400 shares. Therefore, we are going to use the threshold of 1400 shares to create two classes: if the number of shares is greater than 1400, then the article is classified as popular; if the number of shares is less than or equal to 1400, then the article is classified as unpopular in order to formulate a classification problem.

We are going to use multiple linear regression for the first task and binary logistic regression for the second one.

For the ensemble models, we will also fit the two different models(bagged trees and boosted trees) and choose the best one after comparing the misclassification rate.

Finally, we will decide the best two different types of models for our target. 

# Data Study

## Description of the Used Data

Since we are going to predict the popularity of an article, we have chosen *shares* as the response. To start with, for the linear regression model, I am going to fit the model with some selected predictors. 

Similarly, for the ensemble model, I am also going to start to fit two models (bagged tree and  boosted tree) with selected predictors from the train set. After comparing the misclassification rate, the best model can be chosen.

## Data Split

```{r, message = FALSE}
# Load the all libraries and set seed for reproducibility
library(tidyverse)
library(ggplot2)
library(caret)
library(randomForest)
library(tree)
library(gbm)

set.seed(2)

# Read data and remove the first two columns since they are non-predictive.

news_pop <- read_csv('OnlineNewsPopularity.csv')[, -c(1, 2)]

params$weekday

# Just consider 'Monday' at first 
dat<- news_pop %>% select(!starts_with('weekday_is_'), params$weekday)

# Check if the data has any missing values
sum(is.na(dat))

```

Since there is no missing data for this dataset, we can slice the data into the train and the test sets, respectively.

```{r}
# Split the 'Monday' data, 70% of the data for training and the rest for testing
train <- sample(1:nrow(dat), size = nrow(dat) * 0.7)
test <- dplyr::setdiff(1:nrow(dat), train)

TrainDat <- dat[train, ]
TestDat <- dat[test, ]
```


## Data Summarizations

After splicing the data into two sets, we can explore the response variable, *shares* and predictors. 

### Response Variable 

```{r, message = FALSE}
# Histogram of 'shares'
ggplot(data = TrainDat, aes(x = shares)) +
  geom_histogram()

```

Since the histogram of share is highly skewed, we can consider to use the log transformation to obtain a new histogram for shares. 

```{r, message = FALSE}
x <- log(TrainDat$shares)

ggplot(data = TrainDat, aes(x)) +
  geom_histogram() + 
  xlab('Log(shares)')
```

After comparing these two histograms, we decide to use the log(shares) as our response for the multiple linear regression model. 

### Predictor Variables
```{r}
summary(TrainDat)
```

From the output above, we can remove the variable *is_weekend* since it seems to be duplicating days of week; five *LDA* variables due to meaningless values and *kw_min_min, kw_avg_min, kw_min_avg* because of negative values. We can obtain a new train and a test set below.

```{r}
# A new train and a test set
TrainDat <- TrainDat %>% select(!starts_with('LDA_'), -c(is_weekend, kw_min_min, kw_avg_min, kw_min_avg))

TestDat <- TestDat %>% select(!starts_with('LDA_'), -c(is_weekend, kw_min_min, kw_avg_min, kw_min_avg))

head(TrainDat)
head(TestDat)

```

# Modeling

## Linear Regression Fit

### Multiple Linear Regression

We are going to fit the multiple linear regression with selected predictors.

```{r}
# Multiple linear model
fit <- lm(log(shares) ~ . , data = TrainDat)
summary(fit)
```

### Binary Logistic Regression

Use *shares* to create the binary variable : diving the shares into two groups (<1400 and >= 1400) and fit the model with the selected predictors as well.

```{r}
# Convert 'shares' into factor: <= 1400: unpopular, > 1400: popular
TrainDat$shares[TrainDat$shares <= 1400] <- 0
TrainDat$shares[TrainDat$shares > 1400] <- 1

TrainDat$shares <- as.factor(TrainDat$shares)
  
# GLM model
glmFit <- glm(shares~. , data = TrainDat, family = 'binomial')
summary(glmFit)
```

### Linear Regression Model Comparsion

We are going to choose a better linear model from the multiple linear model  and the binary logistic regression model with comparing the values of AIC and BIC respectively.

```{r}
AIC(fit, glmFit)
BIC(fit, glmFit)

```

Since the binary logistic regression model has much smaller AIC and BIC
values, we can choose the binary logistic regression model.

## Ensemble model Fit

For the ensemble model fit, we are going to use the R machine learning `caret` package. 

### Bagged Tree 

```{r, message = FALSE}
# Use ‘trainControl()‘ to control the computational nuances of the train method
trctrl <- trainControl(method = 'repeatedcv', number = 5, repeats = 2)

# Fit a bagged tree
baggedTree <- train(shares ~ ., data = TrainDat, trControl=trctrl,
preProcess = c("center", "scale"), method = "treebag")

baggedTree

# Convert 'shares' in mondayTest into factors as well
TestDat$shares[TestDat$shares <= 1400] <- 0
TestDat$shares[TestDat$shares > 1400] <- 1
TestDat$shares <- as.factor(TestDat$shares)

# Predict classes for test dataset
test_pred_baggedTree <- predict(baggedTree, newdata = TestDat)

# Accurary of the model
confusionMatrix(test_pred_baggedTree, TestDat$shares)
```

### Boosted Tree

```{r, message = FALSE}
# Fit a boosted tree
boostTree <- train(shares ~ ., data = TrainDat, distribution = "bernoulli",
                   method = "gbm", verbose = FALSE, trControl=trctrl)

# Predict classes for test dataset
test_pred_boostTree <- predict(boostTree, newdata = TestDat)

# Accurary of the model
confusionMatrix(test_pred_boostTree, TestDat$shares)

```
### Ensemble model Model Comparsion

We are going to choose a better ensemble model after comparing the misclassification rate.

```{r}

baggedTbl <- table(data.frame(pred = predict(baggedTree, TestDat), true = TestDat$shares))

boostTbl <- table(data.frame(pred = predict(boostTree, TestDat), true = TestDat$shares))

#misclassificatoon rate
1-c(bag = sum(diag(baggedTbl)/sum(baggedTbl)),
    boost = sum(diag(boostTbl)/sum(boostTbl)))
```

Since the boosted trees model with a smaller misclassification rate 0.336 and a larger accuracy rate 0.665 than the bagged trees model, we can choose the boosted trees model.

# Conclusions

For the linear regression model, the binary logistic model is better than the multiple linear regression model after comparing AIC and BIC values. 

For the chosen model, the binary logistic model, we can see that some variables whose coefficients are positive need to be increased while the rest variables whose coefficients are negative needs to be decreased. 

For the ensemble model, the boosted tree model is the best one. This gives highest accuracy around 65%. The data set on a whole gives average accuracy around 64%  which shows that the dataset is inconsistent indicating that irrelevant information has been used.

Therefore, this dataset is insufficient to predict the number of shares with high levels of accuracy for a news article considering its popularity and thus more data needs to be collected.